---
title: "Supplemental Materials for: A robust and readily implementable method for the meta-analysis of response ratios with and without missing standard deviations"
author: Shinichi Nakagawa, Daniel W. A. Noble, Malgorzata Lagisz, Rebecca Spake, Wolfgang Viechtbauer and Alistair M. Senior 
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, tidy = TRUE)
options(digits=2)
```

```{r klippy, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
#install.packages("devtools")
remotes::install_github("rlesur/klippy")
klippy::klippy(tooltip_message = 'Click to Copy Code', tooltip_success = 'Done', position = 'right', color = "red")
```
# Citation
To cite this supplement please use the following paper: 

Shinichi Nakagawa, Daniel W. A. Noble, Malgorzata Lagisz, Rebecca Spake, Wolfgang Viechtbauer and Alistair M. Senior. 2022. A robust and readily implementable method for the meta-analysis of response ratios with and without missing standard deviations. 


# Introduction

In this supplement we show readers how they can implement some of the missing data approaches covered within our paper to deal with missing standard deviations when using log response ratio (*lnRR*) effect size estimates. We provide code snippets that can be copied and make use of a worked example throughout. 
